Synthesis of Linear Functional Programs

### 2021-05-09 article/article.tex

\mypara{Program Synthesis from Polymorfic Refinement Types}~\cite{DBLP:conf/pldi/PolikarpovaKS16}
This paper also attempts to synthetize recursive functional programs, in a more ``advanced'' setting. The specification
is a combination of two rich types, polymorfic and refinement types (which correspond to a first-order logic through the Curry-Howard isomorphism) -- this is an interesting
combo that offers ``expressive power and decidability, which enables automatic verification--and hence synthesis--of nontrivial programs''.
Their approach to the refinement types is a new algorithm that supports decomposition of the refinement specification,
and they separation between the language of specification and programs.

\mypara{Resourceful Program Synthesis from Graded Linear Types}~\cite{DBLP:conf/lopstr/HughesO20}
This paper attempts to synthetize programs in a manner very similar to ours.
It uses graded modal types, which is an ``augment'' to linear types -- a more \emph{fine-grained} version, since it allows for
quantitative specification of resource usage, in contrast to ours either \emph{linear} or \emph{unrestricted} (``banged'' values) use of assumptions.
Their resource management is more complex, and so they provide solutions which adapt Hodas and Miller's input-output model~\cite{} -- the model we'll be using.
They also use focusing as a solution to trim down search space, and our solution will use the same literature as a base.
However, since the logic behind the their types is \emph{modal}, instead of \emph{linear}, it doesn't allow to, in the future, explore
the tight relation between linear logic and concurrency based (message passing) programs with session types (e.g. synthetize concurrent programs with message passing), and other fields such as
resource analysis? -- which ours will.

>>>>

\mypara{Program Synthesis from Polymorfic Refinement Types} The
work~\cite{DBLP:conf/pldi/PolikarpovaKS16} also studies synthesis
recursive functional programs but in a more ``advanced'' setting. The
specification is a combination of two rich forms of types: polymorphic and
refinement types (which correspond to a first-order logic through the
Curry-Howard isomorphism) -- this is an interesting combination that offers
``expressive power and decidability, which enables automatic
verification--and hence synthesis--of nontrivial programs''.  Their
approach to refinement types consists a new algorithm that supports
decomposition of the refinement specification, allowing for separation
between the language of specification and programs and making the
approach amenable to compositional synthesis techniques.

\mypara{Resourceful Program Synthesis from Graded Linear Types} The
work~\cite{DBLP:conf/lopstr/HughesO20} synthesizes programs using an
approach very similar to our own.  The work employs so-called graded
modal types, which is an refinement of pure linear types -- a more
\emph{fine-grained} version, since it allows for quantitative
specification of resource usage, in contrast to ours either
\emph{linear} or \emph{unrestricted} (via the linear logic
exponential) use of assumptions.  Their resource management is more
complex, and so they provide solutions which adapt Hodas and Miller's
resource management approach~\cite{DBLP:journals/tcs/CervesatoHP00,DBLP:journals/tcs/LiangM09} -- the model we'll be using.

They also use focusing as a solution to trim down search space and
to ensure that synthesis only produces well-typed programs, and
our solution will use the same literature as a base.  However, since
their underlying logic is \emph{modal} rather than purely
\emph{linear}, it lacks a clear correspondence with concurrent
session-typed programs~\cite{}, which is a crucial avenue of future
work. Moreover, their use grading effectively requires constraint solving to
be integrated with the synthesis procedure, which can limit the
effectiveness of the overall approach as one scales to more
sophisticated settings (e.g.~refinement types).
